Distributional Trees for Circular Data

### 2019-02-07 pkg/disttree/inst/IWSM2019/abstract_IWSM2019_schlosser.tex
Update fig and include acknowledgment
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2450 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: a3edb39fd4e8a07f484632470769773bdece6407

In this study distributional trees are adapted to circular responses by employing the von Mises
distribution, also known as ``the circular normal distribution''. Based on a location parameter
$\mu \in [-\pi, \pi]$ and a concentration parameter $\kappa > 0$ the density for $y \in [-\pi, \pi]$
is given by:
\begin{equation}
  f_\mathrm{vM}(y; \mu, \kappa) = \frac{1}{2 \pi I_0(\kappa)}~e^{ \kappa \cos(y - \mu)}\label{schlosser:equ_vm}
\end{equation}
where $I_0(\kappa)$ is the modified Bessel function of the first kind and order $0$
(see, e.g., Jammalamadaka and Sengupta (2001) for a more detailed overview).

@@

\begin{figure}[p!]\centering
\includegraphics[height = .5\textheight,angle=90,origin=c]{schlosser-circtree_plot.pdf}
\caption{Fitted tree based on the von Mises distribution for wind direction forecasting.
In each terminal node the empirical histogram (gray) and fitted density (red line)
are depicted along with the estimated location parameter (red hand). The covariates
employed are wind direction ($degree$), wind speed ($\text{ms}^{-1}$),
and pressure gradients ($\text{hPa}$) west and east of the airport,
all lagged by one hour.}
\label{schlosser:fig_tree} \end{figure}

@@

% Acknowledgments, if needed:
\acknowledgments{Special Thanks to ... }

>>>>

In this study distributional trees are adapted to circular responses by employing the von Mises
distribution, also known as ``the circular normal distribution''. Based on a location parameter
$\mu \in [0, 2\pi]$ and a concentration parameter $\kappa > 0$ the density for $y \in [0, 2\pi]$
is given by:
\begin{equation}
  f_\mathrm{vM}(y; \mu, \kappa) = \frac{1}{2 \pi I_0(\kappa)}~e^{ \kappa \cos(y - \mu)}\label{schlosser:equ_vm}
\end{equation}
where $I_0(\kappa)$ is the modified Bessel function of the first kind and order $0$
(see, e.g., Jammalamadaka and Sengupta (2001) for a more detailed overview).

@@

\begin{figure}[p!]\centering
\includegraphics[height = .5\textheight,angle=90,origin=c]{schlosser-circtree_plot.pdf}
\caption{Fitted tree based on the von Mises distribution for wind direction forecasting.
In each terminal node the empirical histogram (gray) and fitted density (red line)
are depicted along with the estimated location parameter (red hand). The covariates
employed are wind direction ($\text{degree}$), wind speed ($\text{ms}^{-1}$),
and pressure gradients ($\text{hPa}$) west and east of the airport,
all lagged by one hour.}
\label{schlosser:fig_tree} \end{figure}

@@

% Acknowledgments, if needed:
\acknowledgments{This project was partially funded by the Austrian Research Promotion 
Agency~(FFG) grant no.~$858537$.}
